Revisiting the Austral Spring Extratropical Southern Hemisphere zonally assymetric circulation using complex Empirical Orthogonal Functions

### 2021-08-17 analysis/paper/shceof.Rmd
Mejora gráficos

@@

```{r test-irlba}
has_irlba <- requireNamespace("irlba", quietly = TRUE)
if (has_irlba) {
  stop("If installed, metR::EOF uses irlba, but that package has a bug with complex matrices\n(see: https://github.com/bwlewis/irlba/issues/55). Uninstall irlba with renv::remove(\"irlba\") and try again.")
}

@@

We apply this method to characterise the Springtime zonally asymmetric circulation of the extratropical Southern Hemisphere.
To take vertical distribution into account, we analyse both the troposphere and the stratosphere.
We focus the analysis to the September-October-November because that's the trimester in which tropical forcing of zonal waves is maximised (cita).
\textcolor{red}{Este parráfo me parece que quedo un poco colgado. Creo que sería importante que esté desde más arriba la idea de la primavera, ya que todo lo otro es muy general, pero al final el análisis solo se va a restringir a esta estación}

@@

    scale_x_longitude(labels = NULL) +
    scale_y_latitude(labels = NULL,
                     minor_breaks = NULL) +
    grid_y +
    facet +
    geom_coords() +
    coord_polar() +
    tag_facets("rc")
}
```

(ref:eof1-regr-gh-cap) Regression of SON geopotential height anomalies (meters) with the (column 1) real and (column 2) imaginary parts of the first cEOF for the 1979 -- 2019 period at (row a) 50 hPa and (row b) 200 hPa. These coefficients come from multiple linear regression involving the real and imaginary parts. Areas marked with dots have p-values smaller than 0.05 adjusted for False Detection Rate.

```{r eof1-regr-gh, fig.cap = "(ref:eof1-regr-gh-cap)", fig.height=1.2*width_column, fig.width=width_column, dependson="plot_regr"}
plot_regr(regr , "cEOF1", c(200, 50), "hgt") +
  scale_fill_divergent("Geopotential height", 
                       breaks = AnchorBreaks(0, exclude = 0),
                       guide = guide_colorsteps_bottom(15, even.steps = TRUE), 
                       limits = c(-1800, 1800))
```

@@

```{r eof2-regr-gh, fig.cap = "(ref:eof2-regr-gh-cap)", fig.height=1.3*width_column, fig.width=width_column, dependson="plot_regr"}
plot_regr(regr, "cEOF2", c(200, 50), "hgt") +
  scale_fill_divergent("Geopotential height", 
                       breaks = AnchorBreaks(0, 200, exclude = 0),
                       guide = guide_colorsteps_bottom(15, even.steps = TRUE), 
                       limits = c(-1000, 1000))
```

@@

(ref:sam-eof-vertical-cap) Coefficient of determination between the real and imaginary part of each cEOF and the SAM, Asymmetric SAM (A-SAM) and Symmetric SAM (S-SAM) indices computed at each level according to @campitelli2021. Thick lines represent estimates with p-value \< 0.01 corrected for False Detection Rate [@benjamini1995].

@@

  scale_x_level("Level (hPa)", minor_breaks = NULL, 
                sec.axis = sa_height_axis(breaks = seq(0, 50, by = 10),
                                          labels = function(x) paste0("~", x, " km"))) +
  scale_y_continuous(r2, limits = c(0, 1)) +
  scale_size_manual("p-value", values = c("TRUE" = 1, 
                                          "FALSE" = 0.5), 
                    guide = "none") +
  scale_color_brewer(NULL, palette = "Dark2", labels = c(full = "SAM",
                                                         asym = "A-SAM",
                                                         sym = "S-SAM")) +
  facet_grid(part ~ cEOF) +
  coord_flip() +
  tag_facets("rc") +
  panel_background +
  theme(panel.spacing = grid::unit(1, "lines"), 
        strip.placement = "outside")
```

@@

```{r eof1-regr-t, fig.cap = "(ref:eof1-regr-t-cap)",  dependson="plot_regr", fig.height=1.3*width_column, fig.width=width_column}
plot_regr(regr, "cEOF1", c(200, 50), "t") +
  scale_fill_divergent("Temperature", breaks = AnchorBreaks(0, exclude = 0),
                       guide = guide_colorsteps_bottom(15, even.steps = TRUE), 
                       limits = c(-4.5, 4.5))
```

@@

```{r t-vertical, fig.cap = "(ref:t-vertical-cap)", fig.height=2.5, fig.width=width_column}
t %>% 
  .[lat %between% o3wave_lats] %>% 
  .[, .(hgt = weighted.mean(hgt, cos(lat*pi/180)),
        o3 = weighted.mean(o3, cos(lat*pi/180)),
        t = weighted.mean(t, cos(lat*pi/180))), by = .(lev, time, lon)] %>%  
  melt(id.vars = c("lev", "time", "lon")) %>% 
  .[, value := Anomaly(value), by = .(time, lev, variable)] %>%
  .[cut(ceof$eof[[1]], 1)$left, on = "time"] %>% 
  sep_ReIm(format = "wider") %>% 
  .[, ":="(Real = as.numeric(scale(Real)), Imaginary = as.numeric(scale(Imaginary))), by = .(cEOF)] %>% 
  .[, FitLm(value, Real, Imaginary, se = TRUE), by = .(lev, lon, variable)] %>% 
  rm_intercept() %>% 
  .[, term := factor_ReIm(term)] %>% 
  ggplot(aes(lon, lev)) +
  geom_contour_fill(data = ~.x[variable == "t"],
                    aes(z = estimate, fill = ..level..),
                    breaks = AnchorBreaks(0, binwidth = 0.3, exclude = 0)) +
  geom_contour_tanaka(data = ~.x[variable == "t"],
                      aes(z = estimate),
                      breaks = AnchorBreaks(0, binwidth = 0.3, exclude = 0)) +
  geom_contour2(data = ~.x[variable == "o3"], 
                aes(z = estimate, linetype = factor(-sign(..level..))), 
                breaks = AnchorBreaks(0, exclude = 0)) +
  geom_text_contour(data = ~.x[variable == "o3"], 
                    aes(z = estimate*1e9, label = signif(..level.., 2)),
                    size = 2,
                    skip = 1, stroke = 0.15, stroke.colour = "white", 
                    label.placement = lable_placement_min(),
                    breaks = AnchorBreaks(0, exclude = 0)) +

  scale_fill_divergent_discretised("Air temperature", 
                                   # limits = c(-2.25, 2.25),
                                   guide = guide_colorsteps_bottom(15)) + 
  scale_y_level(minor_breaks= NULL, 
                sec.axis = sa_height_axis(breaks = seq(0, 50, by = 10),
                                          labels = function(x) paste0("~", x, " km"))) +
  scale_linetype(guide = "none") +
  scale_x_longitude(ticks = 60) +
  facet_grid(~term) +
  axis_labs_smol +
  coord_cartesian(clip = "off") +
  tag_facets()
```

@@

(ref:o3-regr-cap) Regression of SON mean Total Ozone Column anomalies (shaded, Dobson Units) with the (a) real and (b) imaginary parts of the cEOF1 for the 1979 -- 2019 period. On contours, the mean zonal anomaly of Total Ozone Column (negative contours in dashed lines, Dobson Units). Areas marked with dots have p-values smaller than 0.05 adjusted for False Detection Rate.

@@

  geom_contour_pval(aes(z = p.value)) +
  annotate("segment", 
           y  = o3wave_lats,
           yend = o3wave_lats,
           x = 0,
           xend = 360,
           size = 0.3, alpha = 0.8
  ) +
  geom_qmap() +
  scale_fill_divergent_discretised("Total Ozone Column", 
                                   limits = c(-30, 30),  
                                   guide = guide_colorsteps_bottom(15)) + 
  scale_x_longitude(labels = NULL) +
  scale_y_latitude(labels = NULL,
                   minor_breaks = NULL) +
  scale_linetype(guide = "none") +
  geom_coords() +
  coord_polar() +
  facet_grid(cEOF~term) +
  tag_facets("panel")

@@

```{r plot_sst_psi}
plot_sst_psi <- function(which_cEOF, which_lev) {
  one_less <- function(x) {
    n <- length(x)
    if (n < 4) return(x)
    x <- round(x, 5)

  sst_gdata <- regr_sst[cEOF == which_cEOF] %>%
    .[lat %between% c(-90, 10)] %>%
    .[, p_val := Pvaluate(estimate, std.error, df, "fdr"), 
      by = .(cEOF, term)] %>%
    .[, var := "SST"]

@@

  # browser()
  stream_gdata %>% 
    periodic(lon = c(0, 360)) %>%
    ggplot(aes(lon, lat)) +
    geom_contour_fill(aes(z = estimate*1e-7, fill = ..level..),
                      breaks = breaks_stream) +
    geom_contour_pval(aes(z = p_val)) +
    geom_streamline(aes(dx = fx, dy = fy), min.L = 3, L = 10,
                    size = 0.1, res = 2, nx = 22, ny = 12, 
                    arrow.angle = 14, arrow.length = 0.2) +
    geom_qmap(keep = .015) +
    scale_fill_divergent_discretised("Streamfunction (m²/s×1e-7)", 
                                     labels = one_less,
                                     low = scales::muted("red"),
                                     high = scales::muted("blue"),
                                     guide = guide_colorsteps_bottom(15, show.limits = FALSE,
                                                                     order = 99)) +
    ggnewscale::new_scale_fill() +

@@

(ref:sst-psi-2-cap) Regression maps of cEOF2 with SST (K, column a) and streamfunction zonal anomalies ($m^2/s\times10^-7$, shaded) with their corresponding activity wave flux (vectors) (column b). Areas marked with dots have p-values smaller than 0.01 adjusted for FDR.

@@

```{r enso-phase, fig.cap = "ONI plotted against cEOF2 phase for th 1979 -- 2019 period. Years with magnitude of cEOF2 greater or smaller than the 50th percentile are shown as orange circles and green diamomnds respectively. Black line is the fit ONI ~ sin(phase) computed by weighted OLS using the magnitude of the cEOF2 as weights.", fig.width = width_column, fig.height=3}
gdata <- ceof[, eof[[1]]$left, by = .(season)] %>%
  .[cEOF == "cEOF2"] %>% 
  # .[, hgt := Anomaly(hgt)] %>% 
  .[, arg := Arg(hgt)] %>%
  .[, mag := Mod(hgt)] %>%
  .[enso, on = "time"] %>%
  na.omit() %>%
  .[, q := ifelse(mag >= quantile(mag, 0.5), "|cEOF2| > 50%", "|cEOF2| < 50%")]

@@

equation <- latex2exp::TeX(paste0("$\\mathrm{ONI} = ", 
                                  signif(coef(cos_model)[1], 2), "\\sin{(\\mathrm{phase})}$"),
                           output = "character")

@@

```{r sst-psi-1, fig.cap = "(ref:sst-psi-1-cap)", fig.width = 7, fig.height=4, dependson="plot_sst_psi", fig.env = "figure*"}
plot_sst_psi(which_cEOF = "cEOF1", which_lev = 200)

@@

    # scale_fill_divergent() +
    scale_fill_divergent_discretised(NULL, 
                                     # limits = c(-.8, .8),
                                     # oob = scales::squish,
                                     low = "#8E521C",
                                     high = "#00665E",
                                     guide = guide_colorsteps_bottom(15)) +
    # scale_fill_divergent_discretised(NULL, guide = guide_colorsteps_bottom(15)) +
    geom_contour_pval(aes(z = p.value)) +
    geom_qmap(keep = 1) +

@@

(ref:pp-oceania-cap) Regression of SON mean precipitation anomalies in Oceania (mm per day, shaded) with the (row a) real and (row b) imaginary pars of the (column 1) cEOF1 and (column 2) cEOF2. For the 1979 -- 2019. Areas marked with dots have p-values smaller than 0.05 adjusted for False Detection Rate. In black is the West Australia region, defined as the box between `r knitr::combine_words(LonLabel(box_australia$lon))` and `r knitr::combine_words(LatLabel(box_australia$lat))`.

@@

equation <- latex2exp::TeX(paste0("$\\mathrm{Precipitation}_{\\mathrm{West Australia}} = ", 
                                  signif(coef(model)[1], 2), "\\sin{(\\mathrm{phase})} + ", 
                                  signif(coef(model)[2], 2), "\\cos{(\\mathrm{phase})}$"), 
                           output = "character")

@@

```{r pp-america, fig.cap = "(ref:pp-america-cap)", fig.width=width_column, fig.height=width_column*1.1}
plot_pp(data = regr_pp, which_cEOF = c("cEOF1", "cEOF2"), which_continent = "america")  +
  scale_fill_divergent_discretised(NULL, 
                                   limits = c(-1, 1),
                                   # oob = scales::squish,
                                   low = "#8E521C",
                                   high = "#00665E",
                                   guide = guide_colorsteps_bottom(15)) 
# annotate("rect", xmin = min(sesa$lon), xmax = max(sesa$lon), 
#          ymin = min(sesa$lat), ymax = max(sesa$lat),
#          fill = NA, color = "gray50", size = 0.3)
#          
```

@@

On the other hand, over Australia, both the Real and Imaginary phase are associated with significant precipitation anomalies, and we further show that the direction of maximum impact is not aligned with our chosen rotation of cEOF2.
However, this underscores the benefit of using complex EOF, since it would be trivial to rotate it.
\textcolor{red}{creo que esta idea convendría desarrollarla mejor para el que no lo tenga tan claro}

>>>>

one_less <- function(x) {
  n <- length(x)
  if (n < 4) return(x)
  x <- round(x, 5)

AnchorBreaks <- function (anchor = 0, binwidth = NULL, exclude = NULL, bins = 10, sym = FALSE) {
  force(anchor)
  force(binwidth)
  force(exclude)
  force(bins)
  force(sym)
  function(x, binwidth2 = NULL) {
    if (sym) {
      length <- max(abs(x - anchor))
      x <- c(anchor - length, anchor + length)
    }
    if (!is.null(binwidth)) 
      binwidth2 <- binwidth
    if (is.null(binwidth2)) {
      binwidth2 <- diff(pretty(x, bins))[1]
    }
    mult <- ceiling((x[1] - anchor)/binwidth2) - 1L
    start <- anchor + mult * binwidth2
    b <- seq(start, x[2] + binwidth2, binwidth2)
    b[!(b %in% exclude)]
  }
}

@@

```{r test-irlba}
x <- "irlba"
has_irlba <- requireNamespace(x, quietly = TRUE)
if (has_irlba) {
  stop("If installed, metR::EOF uses irlba, but that package has a bug with complex matrices\n(see: https://github.com/bwlewis/irlba/issues/55). Uninstall irlba with renv::remove(\"irlba\") and try again.")
}

@@

We apply this method to characterise the Springtime zonally asymmetric circulation of the extratropical Southern Hemisphere.
To take vertical distribution into account, we analyse both the troposphere and the stratosphere.

@@

    scale_x_longitude(labels = NULL) +
    scale_y_latitude(labels = NULL,
                     minor_breaks = NULL) +
    grid_y +
    facet +
    geom_coords() +
    coord_polar() +
    tag_facets("rc")
}

mgp <- latex2exp::TeX(r'(Geopotential height ($m^2s^{-1}$))')
```

(ref:eof1-regr-gh-cap) Regression of SON geopotential height anomalies ($m^2s^{-1}$) with the (column 1) real and (column 2) imaginary parts of the first cEOF for the 1979 -- 2019 period at (row a) 50 hPa and (row b) 200 hPa. These coefficients come from multiple linear regression involving the real and imaginary parts. Areas marked with dots have p-values smaller than 0.01 adjusted for False Detection Rate.

```{r eof1-regr-gh, fig.cap = "(ref:eof1-regr-gh-cap)", fig.height=1.2*width_column, fig.width=width_column, dependson="plot_regr"}
plot_regr(regr , "cEOF1", c(200, 50), "hgt") +
  scale_fill_divergent(mgp,
                       breaks = AnchorBreaks(0, exclude = 0),
                       guide = guide_colorsteps_bottom(15, even.steps = TRUE), 
                       limits = c(-1800, 1800)) 
```

@@

```{r eof2-regr-gh, fig.cap = "(ref:eof2-regr-gh-cap)", fig.height=1.3*width_column, fig.width=width_column, dependson="plot_regr"}
plot_regr(regr, "cEOF2", c(200, 50), "hgt") +
  scale_fill_divergent(mgp, 
                       breaks = AnchorBreaks(0, 200, exclude = 0),
                       guide = guide_colorsteps_bottom(15, even.steps = TRUE), 
                       limits = c(-1000, 1000))
```

@@

(ref:sam-eof-vertical-cap) Coefficient of determination between the real and imaginary part of each cEOF and the SAM, Asymmetric SAM (A-SAM) and Symmetric SAM (S-SAM) indices computed at each level according to @campitelli2021 for the 1979 -- 2019 period. Thick lines represent estimates with p-value \< 0.01 corrected for False Detection Rate [@benjamini1995].

@@

  scale_x_level("Level (hPa)", minor_breaks = NULL, 
                sec.axis = sa_height_axis(breaks = seq(0, 50, by = 10),
                                          labels = function(x) paste0("~", x, " km"))) +
  scale_y_continuous(r2, limits = c(0, 1), 
                     guide = guide_axis(check.overlap = TRUE)) +
  scale_size_manual("p-value", values = c("TRUE" = 1, 
                                          "FALSE" = 0.5), 
                    guide = "none") +
  scale_color_brewer(NULL, palette = "Dark2", labels = c(full = "SAM",
                                                         asym = "A-SAM",
                                                         sym = "S-SAM")) +
  facet_grid(part ~ cEOF) +
  coord_flip() +
  tag_facets("rc") +
  panel_background +
  theme(panel.spacing = grid::unit(1, "lines"), 
        strip.placement = "outside")
```

@@

```{r eof1-regr-t, fig.cap = "(ref:eof1-regr-t-cap)",  dependson="plot_regr", fig.height=1.3*width_column, fig.width=width_column}
plot_regr(regr, "cEOF1", c(200, 50), "t") +
  scale_fill_divergent("Temperature (K)", breaks = AnchorBreaks(0, exclude = 0),
                       guide = guide_colorsteps_bottom(15, even.steps = TRUE), 
                       limits = c(-4.5, 4.5))
```

@@

```{r t-vertical, fig.cap = "(ref:t-vertical-cap)", fig.height=2.5, fig.width=width_column}
t %>% 
  .[lat %between% o3wave_lats] %>% 
  .[, .(hgt = weighted.mean(hgt, cos(lat*pi/180)),
        o3 = weighted.mean(o3, cos(lat*pi/180)),
        t = weighted.mean(t, cos(lat*pi/180))), by = .(lev, time, lon)] %>%  
  melt(id.vars = c("lev", "time", "lon")) %>% 
  .[, value := Anomaly(value), by = .(time, lev, variable)] %>%
  .[cut(ceof$eof[[1]], 1)$left, on = "time"] %>% 
  sep_ReIm(format = "wider") %>% 
  .[, ":="(Real = as.numeric(scale(Real)), Imaginary = as.numeric(scale(Imaginary))), by = .(cEOF)] %>% 
  .[, FitLm(value, Real, Imaginary, se = TRUE), by = .(lev, lon, variable)] %>% 
  rm_intercept() %>% 
  .[, term := factor_ReIm(term)] %>% 
  ggplot(aes(lon, lev)) +
  geom_contour_fill(data = ~.x[variable == "t"],
                    aes(z = estimate, fill = ..level..),
                    breaks = AnchorBreaks(0, binwidth = 0.25, exclude = 0, sym = TRUE)) +
  geom_contour_tanaka(data = ~.x[variable == "t"],
                      aes(z = estimate),
                      breaks = AnchorBreaks(0, binwidth = 0.25, exclude = 0, sym = TRUE)) +
  geom_contour2(data = ~.x[variable == "o3"], 
                aes(z = estimate, linetype = factor(-sign(..level..))), 
                breaks = AnchorBreaks(0, exclude = 0)) +
  geom_text_contour(data = ~.x[variable == "o3"], 
                    aes(z = estimate*1e9, label = signif(..level.., 2)),
                    size = 2,
                    skip = 1, stroke = 0.15, stroke.colour = "white", 
                    label.placement = lable_placement_min(),
                    breaks = AnchorBreaks(0, exclude = 0)) +

  scale_fill_divergent_discretised("Air temperature (K)", 
                                   # limits = c(-2.25, 2.25),
                                   labels = one_less,
                                   guide = guide_colorsteps_bottom(15, show.limits = TRUE)) + 
  scale_y_level(minor_breaks= NULL, 
                sec.axis = sa_height_axis(breaks = seq(0, 50, by = 10),
                                          labels = function(x) paste0("~", x, " km"))) +
  scale_linetype(guide = "none") +
  scale_x_longitude(ticks = 60) +
  facet_grid(~term) +
  axis_labs_smol +
  coord_cartesian(clip = "off") +
  tag_facets()
```

@@

(ref:o3-regr-cap) Regression of SON mean Total Ozone Column anomalies (shaded, Dobson Units) with the (a) real and (b) imaginary parts of the cEOF1 for the 1979 -- 2019 period. On contours, the mean zonal anomaly of Total Ozone Column (negative contours in dashed lines, Dobson Units). Areas marked with dots have p-values smaller than 0.01 adjusted for False Detection Rate.

@@

  geom_contour_pval(aes(z = p.value)) +
  annotate("segment", 
           y  = o3wave_lats,
           yend = o3wave_lats,
           x = 0,
           xend = 360,
           size = 0.3, alpha = 0.8
  ) +
  geom_qmap() +
  scale_fill_divergent_discretised("Total Ozone Column (DU)", 
                                   limits = c(-30, 30), 
                                   breaks = AnchorBreaks(exclude = 0, binwidth = 5)(c(-30, 30)),
                                   guide = guide_colorsteps_bottom(15)) + 
  scale_x_longitude(labels = NULL) +
  scale_y_latitude(labels = NULL,
                   minor_breaks = NULL) +
  scale_linetype(guide = "none") +
  geom_coords() +
  coord_polar() +
  facet_grid(cEOF~term) +
  tag_facets("panel")

@@

```{r plot_sst_psi}
plot_sst_psi <- function(which_cEOF, which_lev) {
  sst_gdata <- regr_sst[cEOF == which_cEOF] %>%
    .[lat %between% c(-90, 10)] %>%
    .[, p_val := Pvaluate(estimate, std.error, df, "fdr"), 
      by = .(cEOF, term)] %>%
    .[, var := "SST"]

@@

  # browser()
  stream_gdata %>% 
    periodic(lon = c(0, 360)) %>%
    ggplot(aes(lon, lat)) +
    geom_contour_fill(aes(z = estimate*1e-7, fill = ..level..),
                      breaks = breaks_stream) +
    geom_contour_pval(aes(z = p_val)) +
    geom_streamline(aes(dx = fx, dy = fy), min.L = 3, L = 10,
                    size = 0.1, res = 2, nx = 22, ny = 12, 
                    arrow.angle = 14, arrow.length = 0.2) +
    geom_qmap(keep = .015) +
    scale_fill_divergent_discretised(latex2exp::TeX(r'(Streamfunction ($m^2s^{-1}\times 1e^{-7}$))'), 
                                     labels = one_less,
                                     low = scales::muted("red"),
                                     high = scales::muted("blue"),
                                     guide = guide_colorsteps_bottom(15, show.limits = FALSE,
                                                                     order = 99)) +
    ggnewscale::new_scale_fill() +

@@

(ref:sst-psi-2-cap) Regression maps of cEOF2 with SST (K, column a) and streamfunction zonal anomalies ($m^2/s\times10^-7$, shaded) with their corresponding activity wave flux (vectors) (column b) for the 1979 -- 2019 period. Areas marked with dots have p-values smaller than 0.01 adjusted for FDR.

@@

(ref:enso-phase-cap) ONI plotted against cEOF2 phase for th 1979 -- 2019 period.
Years with magnitude of cEOF2 greater or smaller than the 50th percentile are shown as orange circles and green diamomnds respectively. 
Black line is the fit ONI ~ sin(phase) computed by weighted OLS using the magnitude of the cEOF2 as weights.

```{r enso-phase, fig.cap = "(ref:enso-phase-cap)", fig.width = width_column, fig.height=3}
gdata <- ceof[, eof[[1]]$left, by = .(season)] %>%
  .[cEOF == "cEOF2"] %>% 
  # .[, hgt := Anomaly(hgt)] %>% 
  .[, arg := Arg(hgt)] %>%
  .[, mag := Mod(hgt)] %>%
  .[enso, on = "time"] %>%
  na.omit() %>%
  .[, q := ifelse(mag >= quantile(mag, 0.5), "|cEOF2| > 50%", "|cEOF2| < 50%")]

@@

equation <- latex2exp::TeX(paste0("$\\mathrm{ONI} = ", 
                                  signif(coef(cos_model)[1], 2), "\\sin{(\\mathrm{phase})}$"))

@@

```{r sst-psi-1, fig.cap = "(ref:sst-psi-1-cap)", fig.width = 7, fig.height=4, dependson="plot_sst_psi", fig.env = "figure*"}
plot_sst_psi(which_cEOF = "cEOF1", which_lev = 200) 
```

@@

    # scale_fill_divergent() +
    scale_fill_divergent_discretised("Precipitation (mm/day)", 
                                     # limits = c(-.8, .8),
                                     # oob = scales::squish,
                                     low = "#8E521C",
                                     high = "#00665E",
                                     guide = guide_colorsteps_bottom(15)) +
    # scale_fill_divergent_discretised(NULL, guide = guide_colorsteps_bottom(15)) +
    geom_contour_pval(aes(z = p.value)) +
    geom_qmap(keep = 1) +

@@

(ref:pp-oceania-cap) Regression of SON mean precipitation anomalies in Oceania (mm per day, shaded) with the (row a) real and (row b) imaginary pars of the (column 1) cEOF1 and (column 2) cEOF2. For the 1979 -- 2019. Areas marked with dots have p-values smaller than 0.01 adjusted for False Detection Rate. In black is the West Australia region, defined as the box between `r knitr::combine_words(LonLabel(box_australia$lon))` and `r knitr::combine_words(LatLabel(box_australia$lat))`.

@@

equation <- latex2exp::TeX(paste0("$\\mathrm{Precipitation}_{\\mathrm{West Australia}} = ", 
                                  signif(coef(model)[1], 2), "\\sin{(\\mathrm{phase})} + ", 
                                  signif(coef(model)[2], 2), "\\cos{(\\mathrm{phase})}$"))

@@

```{r pp-america, fig.cap = "(ref:pp-america-cap)", fig.width=width_column, fig.height=width_column*1.1}
plot_pp(data = regr_pp, which_cEOF = c("cEOF1", "cEOF2"), which_continent = "america")  +
  scale_fill_divergent_discretised("Precipitation (mm/day)", 
                                   limits = c(-1, 1),
                                   # oob = scales::squish,
                                   low = "#8E521C",
                                   high = "#00665E",
                                   guide = guide_colorsteps_bottom(15)) 
# annotate("rect", xmin = min(sesa$lon), xmax = max(sesa$lon), 
#          ymin = min(sesa$lat), ymax = max(sesa$lat),
#          fill = NA, color = "gray50", size = 0.3)
#          
```

@@

On the other hand, over Australia, both the Real and Imaginary phase are associated with significant precipitation anomalies, and we further show that the direction of maximum impact is not aligned with our chosen rotation of cEOF2.
However, this underscores the benefit of using complex EOF, since it would be trivial to rotate it.
That is, look for the rotation of complex EOF that maximises the relationship and use that as the relevant index for study or forecast.
